Compositional equivalences based on Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-10-21 2016-Weak-FH-Bisim/WeakBisim.tex
typo in OT Def
related bib

- TROUVER UN EXEMPLE ET LE METTRE DANS L ARTICLE

@@

- adopter la def semantique d ot pour la bisim

@@

Fig. \ref{SimpleProt:Schema} shows the example principle: two unspecified
processes $P$ and $Q$ (Holes) communicate messages, with a data value
argument, through the two protocol entities $sender$ and
$receiver$. The underlying network is modelled by a medium entity transporting messages from the sender to the receiver, and that is able to detect some transport error, and signal it to the sender. The return $ack$ message from $receiver$ to $sender$ is supposed to be safe. The final transmission of the message to the recipient (Hole Q) include the value of the ``error counter'' $ec$.

@@

Fig. \ref{SimpleProt:Spec} shows the Schema of this system, and on the right side, its global specification, expressed as a pNet. 
The full system should be equivalent (e.g. through weak bisimulation)
to the two processes connected simply through a perfect medium. Fig. \ref{SimpleProt:Spec}, right, shows the pNet representing this global specification.

@@

\subsection{Open Automata}
 Open automata (OA) are not composition structures but they are made of transitions that are dependent of the actions of the holes, and they can reason on variables (potentially with only symbolic values). Formaly, each state of an OA has set of \emph{state variables}, and all state variable sets are disjoint.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]\label{def:OT}
	\label{def:OpenTransitions}
	An \emph{open transition} (OT) over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$ and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\openrule
	{	\beta_j^{j\in J'}, \Pred, \Post}
	{s \OTarrow {\alpha}s'}
	\end{mathpar}
	Where $J'\subseteq$, $s, s'\in\mathcal{S}$ and $\beta_j$
        is a transition of the hole $j$, with $\beta_j\in\Sort_j$. $\alpha$ is an action 
        label denoting the resulting action of this open transition.
        We call \emph{source variables} of an OT the union of all variables
        in the different terms
        $\beta_j$ and $\alpha$, and the state variables of $s$.
        \Pred\ is a predicate over the source variables. \Post\ is a set of 
	equations that hold \emph{after the open transition}, they are
        represented as a substitution of the form $({x_k\gets e_k})^{k\in K}$ 
	where $x_k$ are state variables of $s'$, and $e_k$ are expressions
        over the source variables. Open transitions are identified
        modulo logical equivalence on their predicate. 
\end{definition}

@@

The proof of this theorem can be found in Appendix~\ref{thm-equiv-proof}. The only non-trivial part of the proof is the proof of transitivity. It relies on the following elements. First,  the transitive composition of two relations with predicate is defined; this is not exactly standard as it necessitates to define the right predicate for the transitive composition and producing a single predicate to relate any two states. Then the fact that one open transition is simulated by a family of open transitions leads to a doubly indexed family of simulating open transition; this needs particular care, also because of the use of renaming (\Post) when proving that the predicates satisfy the definition (property on $\Pred_{s,t} \land \Pred_{OT}$ in the definition).

@@

\TODO{rertavailler le paragraphe ci dessous}
In
\cite{Johnson:2013:CBSE,Johnson:2014:qosa}, the authors investigate several methodologies for the compositional verification of software systems, with the aim to verify reconfigurable component systems. To improve scaling and compositionality, the authors  decouple the verification problem that is to be resolved by a SMT (satisfiability modulo theory) solver into independent sub-problems on independent sets of variables. These works clearly highlight the interest of incremental and compositional verification in a very general setting. By adding more structure to the composition model and designing a dedicated bisimulation theory, we show here how to enforce a compositional proof system that is more powerful than independent sets of variables. Our theory has also been encoded into an SMT solver and it would be interesting to investigate how the examples of evolving systems studied by the authors could be encoded into pNet and verified by our framework. However, the models of Johnson et al. are quite different from ours, in particular they are much less structured, and translating them is clearly outside the scope of this article. Additionally, the more compositional constraint resolution provided by the authors could also be interesting to make our SMT proofs scale better (inside a single pNet for example), the integration of the Z3 framework described in \cite{Johnson:2014:qosa} with our proof theory for open pNets would also be worth investigating from a practical point of view. One major different however is that pNets in their current state are not much adapted to probabilistic verification contrarily to \cite{Johnson:2014:qosa} but they are better suited to verify complex behavioural properties.

@@

\NOTE{Eric: one transition of the <020> table, from the draft on my white board}

>>>>

@@

@@

Fig. \ref{SimpleProt:Spec} (left) shows the example principle: two unspecified
processes $P$ and $Q$ (Holes) communicate messages, with a data value
argument, through the two protocol entities $sender$ and
$receiver$. The underlying network is modelled by a medium entity transporting messages from the sender to the receiver, and that is able to detect some transport error, and signal it to the sender. The return $ack$ message from $receiver$ to $sender$ is supposed to be safe. The final transmission of the message to the recipient (Hole Q) include the value of the ``error counter'' $ec$.

@@

Fig. \ref{SimpleProt:Spec} (right) shows the Schema of this system, and on the right side, its global specification, expressed as a pNet. 
The full system should be equivalent (e.g. through weak bisimulation)
to the two processes connected simply through a perfect medium. Fig. \ref{SimpleProt:Spec}, right, shows the pNet representing this global specification.

@@

\subsection{Open Automata}
 Open automata (OA) are not composition structures but they are made of transitions that are dependent of the actions of the holes, and they can reason on variables (potentially with only symbolic values). Formaly, each state of an OA has set of \emph{state variables}, and all state variable sets are disjoint.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]\label{def:OT}
	\label{def:OpenTransitions}
	An \emph{open transition} (OT) over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$ and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\openrule
	{	\beta_j^{j\in J'}, \Pred, \Post}
	{s \OTarrow {\alpha}s'}
	\end{mathpar}
	Where $J'\subseteq J$, $s, s'\in\mathcal{S}$ and $\beta_j$
        is a transition of the hole $j$, with $\beta_j\in\Sort_j$. $\alpha$ is an action 
        label denoting the resulting action of this open transition.
        We call \emph{source variables} of an OT the union of all variables
        in the different terms
        $\beta_j$ and $\alpha$, and the state variables of $s$.
        \Pred\ is a predicate over the source variables. \Post\ is a set of 
	equations that hold \emph{after the open transition}, they are
        represented as a substitution of the form $({x_k\gets e_k})^{k\in K}$ 
	where $x_k$ are state variables of $s'$, and $e_k$ are expressions
        over the source variables. Open transitions are identified
        modulo logical equivalence on their predicate. 
\end{definition}

@@

The proof of this theorem can be found in~\ref{thm-equiv-proof}. The only non-trivial part of the proof is the proof of transitivity. It relies on the following elements. First,  the transitive composition of two relations with predicate is defined; this is not exactly standard as it necessitates to define the right predicate for the transitive composition and producing a single predicate to relate any two states. Then the fact that one open transition is simulated by a family of open transitions leads to a doubly indexed family of simulating open transition; this needs particular care, also because of the use of renaming (\Post) when proving that the predicates satisfy the definition (property on $\Pred_{s,t} \land \Pred_{OT}$ in the definition).

@@

\TODO{retravailler le paragraphe ci dessous}
In
\cite{Johnson:2013:CBSE,Johnson:2014:qosa}, the authors investigate several methodologies for the compositional verification of software systems, with the aim to verify reconfigurable component systems. To improve scaling and compositionality, the authors  decouple the verification problem that is to be resolved by a SMT (satisfiability modulo theory) solver into independent sub-problems on independent sets of variables. These works clearly highlight the interest of incremental and compositional verification in a very general setting. By adding more structure to the composition model and designing a dedicated bisimulation theory, we show here how to enforce a compositional proof system that is more powerful than independent sets of variables. Our theory has also been encoded into an SMT solver and it would be interesting to investigate how the examples of evolving systems studied by the authors could be encoded into pNet and verified by our framework. However, the models of Johnson et al. are quite different from ours, in particular they are much less structured, and translating them is clearly outside the scope of this article. Additionally, the more compositional constraint resolution provided by the authors could also be interesting to make our SMT proofs scale better (inside a single pNet for example), the integration of the Z3 framework described in \cite{Johnson:2014:qosa} with our proof theory for open pNets would also be worth investigating from a practical point of view. One major different however is that pNets in their current state are not much adapted to probabilistic verification contrarily to \cite{Johnson:2014:qosa} but they are better suited to verify complex behavioural properties.

\ERIC{J'ai copie/colle ici le Related Work que Zechen a mis dans notre
  papier bisim...
  Pas tres satisfaisant pour les aspects Weak, mais quelques inputs a
  regarder}

To the best of our knowledge, there is no other research works on the Bisimulation Equivalence between such complicate(open, symbolic, parameterized, with loops and assignments) system models, with providing fully complete algorithms.
For the sake of completeness, we give a brief overview of other Symbolic Bisimulation researches.

One line of research is providing Symbolic Bisimulation for different model or language. 
In Calder's work\cite{calder2001symbolic}, they define a symbolic semantic for full LOTOS, with a symbolic bisimulation over it;
Borgstrom et al., Delaune et al. and Buscemi et al. provide symbolic semantic and equivalence for different variants of pi calculus respectively\cite{borgstrom2004symbolic}\cite{delaune2007symbolic}\cite{buscemi2008open}, and later in 2012 Liu's work provided symbolic bisimulation  for full applied pi calculus\cite{liu2010complete};
The most recent work, Feng et al. provide a symbolic bisimulation for quantum processes
\cite{feng2014symbolic}.
All the above works, didn't give a complete approach for verification, and the models on which these works based are simpler or different than ours.

Another line of research is devising the algorithm for computing symbolic Bisimulation Equivalence.
In Dovier's work, they provide a rank-based algorithm, layering the input model to compute bisimulation\cite{dovier2002rank}. 
Baldan et al. also focus on open systems, using a logical programming language \emph{Prolog} to model the systems and compute Bisimulation\cite{baldan2001compositional}.
Wimmer et al. present a signature-based approach to compute Bisimulation, implemented by using BDDs\cite{wimmer2006sigref}.
Lin present a symbolic bisimulation between symbolic transition graph with assignments(STGA), as mentioned in \emph{Introduction}, bring us lots of inspirations.

@@

\NOTE{ERIC: one transition of the <020> table, from the draft on my white board}
